The difference operation between templates of binary cellular automata

### 2015-11-17 article/article.tex
Elsevier format

\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[brazil]{babel}
\usepackage{graphicx} \graphicspath{ {../img/} }
\usepackage{hyperref}
\usepackage{abnt-alf}
\usepackage[top=3cm,bottom=2cm,left=3cm,right=2cm]{geometry}
\usepackage{indentfirst}
\usepackage[table,xcdraw]{xcolor}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{import}

% O tamanho do parágrafo é dado por:
\setlength{\parindent}{1.3cm}

% Controle do espaçamento entre um parágrafo e outro:
\setlength{\parskip}{0.2cm}  % tente também \onelineskip
% Retira espaço extra obsoleto entre as frases.
\frenchspacing

\title{The difference operation between templates of binary cellular automata}
\author{Zorandir Soares Junior \footnote{zorandir@gmail.com} \\
 	Programa de Pós-Graduação em Engenharia Elétrica e Computação \\
	Universidade Presbiteriana Mackenzie
	\and
	Pedro Paulo Balbi de Oliveira \footnote{pedrob@mackenzie.br}  \\
 	Programa de Pós-Graduação em Engenharia Elétrica e Computação \\
	Universidade Presbiteriana Mackenzie
	\and
	Maurício Verardo \footnote{mverardo@gmail.com}  \\
 	Programa de Pós-Graduação em Engenharia Elétrica e Computação \\
	Universidade Presbiteriana Mackenzie
	}

\date{\today}

\begin{document}
\citeoption{abnt-etal-cite=3}
\citeoption{abnt-etal-list=5}
\citeoption{abnt-etal-text=3}

% DESENVOLVIMENTO
\pagestyle{plain}
\renewcommand{\baselinestretch}{1.25} 
\normalsize

\maketitle

\begin{abstract}
 The notion of templates has recently been introduced to directly represent families of cellular automata that share certain property. Here we introduce the difference operation between templates, whose applicability is in representing families of binary cellular automata that share some properties but not others. This operation requires another process that generates what we call the \textit{exception templates}, which are also introduced here. The use of both defined operations is illustrated with examples in the space of elementary cellular automata.

\begin{flushleft}
{\bf Palavras-chave:} {\it Cellular automata, rule space, templates, difference of templates, exception of templates.}
\end{flushleft}
\end{abstract}

\section{Introduction}
\label{sec:introducao}
Cellular automata (CAs) are dynamical systems typically discrete in time, space and states.
Through rules of simple local action, CAs can produce behavior of great complexity. \cite{wolfram2002}. The study of classical problems of cellular automata, as the problem of parity \cite{Betel2013} and the density problem \cite{deOliveira2014density}, can help in the understanding of how this complex behavior emerges. %TODO Mudei de local a citação wolfram2002

@@

After the next section presents some basics about cellular automata, Section \ref{sec:templates} presents in more detail what are templates and its major operations. The Section \ref{sec:diferenca_entre_templates_e_templates_de_excecao} introduces the operation of difference between templates and the operation that generates what we call the \textit{exception templates}, as well as their applications. Finally, the Section \ref{sec:consideracoes_finais} makes closing remarks about the news algorithms and mentions future work.

@@

According to Boccara and Fukś (\citeyear{boccara2002}), an AC is conservative when each of its local function $f$, applied in every neighborhood $(\alpha_0,\alpha_1, \dots, \alpha_{n-1})$ satisfy the conditions described in Eq. \eqref{eq:conservativeCA}.
\begin{equation}
\begin{split}
f(\alpha_0,\alpha_1, \dots,\alpha_{n-1}) = \alpha_0 + (\sum_{i=0}^{n-2}f(0_0,0_1, \dots,0_i,\alpha_1,\alpha_2, \dots,\alpha_{n-1}) \\- f(0_0,0_1, \dots,0_i,\alpha_0,\alpha_1, \dots,\alpha_{n-i-1}))
\label{eq:conservativeCA}
\end{split}
\end{equation}

@@

	\begin{figure}[h!]
	  \centering
	  \includegraphics[width=.5\textwidth]{fig_ruleIcon60.png}
	  \caption{Tabela de transições da regra 60 do espaço elementar.}
	  \label{fig:table60}
	\end{figure}

	\begin{figure}[h!]
	  \centering
	  \includegraphics[width=.5\textwidth]{fig_ruleIcon102.png}
	  \caption{Tabela de transições da regra 102 do espaço elementar, obtida através da transformação de reflexão aplicada na tabela de transições da regra 60.}
	  \label{fig:table102}
	\end{figure}

@@

\section{Templates}
\label{sec:templates}
Template is a generalization of the state transition tables of CAs through variables. A single template may represent a rule set that may present particular property. The templates were created and implemented by De Oliveira and Verardo (\citeyear{deOliveira2014}) and are currently available at the library \textit{open source CATemplates} \cite{CATemplates} on GitHub.

@@

Besides the possibility of expansion of templates, it has been developed by De Oliveira e Verardo (\citeyear{deOliveira2014b}) an algorithm to generate templates that represent the intersection between two sub-spaces of rules represented by templates. To illustrate, consider the search for a rule set that are both color blind and have maximum internal symmetry by composition. To find this rule set is necessary to perform two steps: first one must match the two templates that represent the desired properties, thus forming an equation system; then to solve the equations, relationships are generated between the variables that, when applied to templates received as input will result in the template representing the intersection.

@@

Finally, apply a negation operation on the equations that, in the binary case, corresponding to the permutation $\rho $ or, equivalently, the implementation of the function $f(x) = 1 - (x)$. Eq. \eqref{eq:differenceP4} represents the logical combination of equations resulting from these operations.
\begin{equation}
1 - x_0	= 1 - 1					\vee 
1 - x_3	= 1 - (1 - x_1 - x_2)	\vee
x_3		= 1 - (x_1 + x_2)		\vee 
x_0		= 1 - 0				
\label{eq:differenceP4}
\end{equation}

@@

\def\refname{Bibliography}
\bibliography{bibliografia}
\addcontentsline{toc}{section}{Bibliography} 
\bibliographystyle{abnt-alf}

\end{document}

>>>>

\documentclass{llncs}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx} \graphicspath{ {../img/} }

\begin{document}
\pagestyle{headings}  % switches on printing of running heads
\addtocmark{The difference operation between templates of binary cellular automata} % additional mark in the TOC

\title{The difference operation between templates of binary cellular automata}

\titlerunning{Difference operation between templates}

\author{Zorandir Soares\inst{2} \and Maurício Verardo\inst{2} \and
Pedro Paulo Balbi de Oliveira\inst{1}\inst{2}}

\authorrunning{Zorandir Soares, Maurício Verardo, and Pedro de Oliveira}

\tocauthor{Zorandir Soares, Maurício Verardo, and Pedro de Oliveira}

\institute{Universidade Presbiteriana Mackenzie\\
Faculdade de Computação e Informática
\and
Pós-Graduação em Engenharia Elétrica e Computação\\
Rua da Consolação 896, Consolação\\
01302-907 São Paulo, SP - Brazil\\
\email{zorandir@gmail.com}\\
\email{mauricio.verardo@gmail.com}\\
\email{pedrob@mackenzie.br}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The notion of templates has recently been introduced to directly represent families of cellular automata that share certain property. Here we introduce the difference operation between templates, whose applicability is in representing families of binary cellular automata that share some properties but not others. This operation requires another process that generates what we call the \textit{exception templates}, which are also introduced here. The use of both defined operations is illustrated with examples in the space of elementary cellular automata.
\keywords{cellular automata, rule space, templates, difference of templates, exception templates.}
\end{abstract}
%
\section{Introduction}
\label{sec:introducao}
Cellular automata (CAs) are dynamical systems typically discrete in time, space and states.
Through rules of simple local action, CAs can produce behavior of great complexity. \cite{wolfram2002}. The study of classical problems of cellular automata, as the problem of parity \cite{Betel2013} and the density problem \cite{deOliveira2014density}, can help in the understanding of how this complex behavior emerges. %TODO Mudei de local a citação wolfram2002

@@

After the next section presents some basics about cellular automata, Section \ref{sec:templates} presents in more detail what are templates and its major operations. The Section \ref{sec:diferenca_entre_templates_e_templates_de_excecao} introduces the operation of difference between templates and the operation that generates what we call the \textit{exception templates}, as well as their applications. Finally, the Section \ref{sec:consideracoes_finais} makes closing remarks about the news algori

@@

According to Boccara and Fukś (\cite{boccara2002}), an AC is conservative when each of its local function $f$, applied in every neighborhood $(\alpha_0,\alpha_1, \dots, \alpha_{n-1})$ satisfy the conditions described in Eq. \ref{eq:conservativeCA}.
\begin{equation}
\begin{split}
f(\alpha_0,\alpha_1, \dots,\alpha_{n-1}) = \alpha_0 + (\sum_{i=0}^{n-2}f(0_0,0_1, \dots,0_i,\alpha_1,\alpha_2, \dots,\alpha_{n-1}) \\- f(0_0,0_1, \dots,0_i,\alpha_0,\alpha_1, \dots,\alpha_{n-i-1}))
\label{eq:conservativeCA}
\end{split}
\end{equation}

@@

  \begin{figure}[h!]
    \centering
    \includegraphics[width=.5\textwidth]{fig_ruleIcon60.png}
    \caption{Transitions table of the rule 60 of the elementary space.}
    \label{fig:table60}
  \end{figure}

  \begin{figure}[h!]
    \centering
    \includegraphics[width=.5\textwidth]{fig_ruleIcon102.png}
    \caption{Transitions table of the rule 102 from the elementary space, obtained through transformation by reflection applied to the transitions table of rule 60.}
    \label{fig:table102}
  \end{figure}

@@

\section{Templates}
\label{sec:templates}
Template is a generalization of the state transition tables of CAs through variables. A single template may represent a rule set that may present particular property. The templates were created and implemented by De Oliveira and Verardo (2014) and are currently available at the library \textit{open source CATemplates} \cite{CATemplates} on GitHub.

@@

Besides the possibility of expansion of templates, it has been developed by De Oliveira e Verardo (2014) an algorithm to generate templates that represent the intersection between two sub-spaces of rules represented by templates. To illustrate, consider the search for a rule set that are both color blind and have maximum internal symmetry by composition. To find this rule set is necessary to perform two steps: first one must match the two templates that represent the desired properties, thus forming an equation system; then to solve the equations, relationships are generated between the variables that, when applied to templates received as input will result in the template representing the intersection.

@@

Finally, apply a negation operation on the equations that, in the binary case, corresponding to the permutation $\rho $ or, equivalently, the implementation of the function $f(x) = 1 - (x)$. Eq. \ref{eq:differenceP4} represents the logical combination of equations resulting from these operations.
\begin{equation}
1 - x_0 = 1 - 1         \vee 
1 - x_3 = 1 - (1 - x_1 - x_2) \vee
x_3   = 1 - (x_1 + x_2)   \vee 
x_0   = 1 - 0       
\label{eq:differenceP4}
\end{equation}

@@

%
% ---- Bibliography ----
%
\begin{thebibliography}{}
\providecommand{\abntrefinfo}[3]{}
\providecommand{\abntbstabout}[1]{}
\abntbstabout{1.52 }

\bibitem[Betel, {DE OLIVEIRA} e Flocchini 2013]{Betel2013}
\abntrefinfo{Betel, {DE OLIVEIRA} e Flocchini}{BETEL; {DE OLIVEIRA};
  FLOCCHINI}{2013}
{BETEL, H.; {DE OLIVEIRA}, P. P.~B.; FLOCCHINI, P. Solving the parity problem
  in one-dimensional cellular automata.
\emph{Natural Computing}, v.~12, n.~3, p. 323--337, 2013.}

\bibitem[Boccara e {FUK\'{S}} 2002]{boccara2002}
\abntrefinfo{Boccara e {FUK\'{S}}}{BOCCARA; {FUK\'{S}}}{2002}
{BOCCARA, N.; {FUK\'{S}}, H. Number-conserving cellular automaton rules.
\emph{Fundamenta Informaticae}, IOS Press, v.~52, n.~1-3, p. 1--13, 2002.}

\bibitem[{DE OLIVEIRA} 2014]{deOliveira2014density}
\abntrefinfo{{DE OLIVEIRA}}{{DE OLIVEIRA}}{2014}
{{DE OLIVEIRA}, P. P.~B. On density determination with cellular automata:
  Results, constructions and directions.
\emph{Journal of Cellular Automata}, v.~9, n.~5-6, p. 357--385, 2014.}

\bibitem[{DE OLIVEIRA} e Verardo 2014]{deOliveira2014}
\abntrefinfo{{DE OLIVEIRA} e Verardo}{{DE OLIVEIRA}; VERARDO}{2014a}
{{DE OLIVEIRA}, P. P.~B.; VERARDO, M. Representing families of cellular
  automata rules.
\emph{The Mathematica Journal}, v.~16, n.~8, 2014.
Dispon{\'\i}vel em: $<$dx\-.doi\-.org\-/doi:10\-.3888\-/tmj\-.16-8$>$.}

\bibitem[{DE OLIVEIRA} e Verardo 2014]{deOliveira2014b}
\abntrefinfo{{DE OLIVEIRA} e Verardo}{{DE OLIVEIRA}; VERARDO}{2014b}
{{DE OLIVEIRA}, P. P.~B.; VERARDO, M. Template based representation of cellular
  automata rules. In:  ISOKAWA, T.; IMAI, K.; MATSIU, N.; PEPER, F.; UMEO, H.
  (Ed.). \emph{20th International Workshop on Cellular Automata and Discrete
  Complex Systems}. Himeji, Japão, Julho 7-9: [s.n.], 2014. p. 199--204.}

\bibitem[Salo e T{\"o}rm{\"a} 2013]{salo2013color}
\abntrefinfo{Salo e T{\"o}rm{\"a}}{SALO; T{\"O}RM{\"A}}{2013}
{SALO, V.; T{\"O}RM{\"A}, I. Color blind cellular automata. In:  \emph{Cellular
  Automata and Discrete Complex Systems}. [S.l.]: Springer, 2013. p. 139--154.}

\bibitem[Verardo e {DE OLIVEIRA} 2015]{CATemplates}
\abntrefinfo{Verardo e {DE OLIVEIRA}}{VERARDO; {DE OLIVEIRA}}{2015}
{VERARDO, M.; {DE OLIVEIRA}, P. P.~B. \emph{CATemplates}.
[S.l.], 2015. Dispon{\'\i}vel em:
  $<$https:\-/\-/github\-.com\-/mverardo\-/CATemplates$>$.}

\bibitem[Wolfram 1994]{wolfram1994cellular}
\abntrefinfo{Wolfram}{WOLFRAM}{1994}
{WOLFRAM, S. \emph{Cellular automata and complexity: collected papers}. [S.l.]:
  Addison-Wesley Reading, 1994.}

\bibitem[Wolfram 2002]{wolfram2002}
WOLFRAM 2002
{WOLFRAM, S. \emph{A new kind of science}. [S.l.]: Wolfram media Champaign,
  2002.}

\bibitem[Wolz e {DE OLIVEIRA} 2008]{wolz2008very}
\abntrefinfo{Wolz e {DE OLIVEIRA}}{WOLZ; {DE OLIVEIRA}}{2008}
{WOLZ, D.; {DE OLIVEIRA}, P. P.~B. Very effective evolutionary techniques for
  searching cellular automata rule spaces.
\emph{J. Cellular Automata}, v.~3, n.~4, p. 289--312, 2008.}

\end{thebibliography}
\end{document}
